Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-04-28 Manuscript/PLoSONEresponse/response.tex

See our detailed response below.

\textcolor{blue}{3. Does the manuscript adhere to the PLOS Data Policy?}

\textcolor{blue}{Reviewer $\#$ 1: No}

We apologise for the confusion here. We mentioned where our data was stored in the supporting information but forgot to add this under the ``Data Reproducibility and Availability" section in the manuscript. We have now corrected this. All the data generated by the simulations are directly available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} (goo.gl/h5Kxb5). Note that along with the data we share the full code used for repeating the submitted manuscript, from simulating the data to generating the figures and compiling the manuscript (available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub} (goo.gl/ewRsrm). All this code is provided with a detailed description of each step that we are happy to develop if needed.

@@

\textcolor{blue}{5. Review Comments to the Author}

\textcolor{blue}{Reviewer $\#$ 1: This paper tackles an interesting questions. Despite some problems noted below, the experimental design seems adequate.}

%-------------

\textcolor{blue}{A-Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions. I suspect that if the authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective), they'll be able to make a statistical case for most of the points that they are trying to make. But I just don't think that a simulation comparison should be publishable without *some* statistical analysis (beyond the calculation of summaries).}

@@

% Talk to Andrew

%We feel that the reviewer has slightly misunderstood the term ``statistics'' in this case. Firstly, p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e. significant ones) when the sample sizes are big; (2) concerning the confidence intervals, we actually did systematically calculate them; ans (3) we are not convinced by a Bayesian approach for comparing distributions.

%-------------

\textcolor{blue}{B-The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}

@@

% Priors comment??? Move stuff up from below

The reviewer is (I believe) referring to their comment below (number 13). We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only. The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of move probability% check what you call it in paper
 for the three priors).

%Our results did indeed biased us in favour of using a Bayesian approach to fix topology since our results clearly indicate that it is less likely to generate a wrong topology using a Bayesian approach (see Figure 4 and 5 and discussion line 472 to 496 in the first submission manuscript).
% TG: or I'm not sure I understood that one.
%-------------
% add Mk = Mkv

\textcolor{blue}{C-The use of Mk rather than Mkv makes the study less relevant to real data analyses.}

Throughout our paper, we referred the M\textit{k} model as being the one described in Lewis (2001) (an equal rate Q-matrix with \textit{k} character states). The Mkv model is identical to the Mk model (as described in Lewis 2001) however, in practice, it assumes that all characters are variable and that therefore, one should correct for acquisition bias that can induce branch length inflation. This is generally by adding an algorithm in the phylogenetic software to add ``dummy'' morphological characters. In some software (e.g. GARLI or BEAST), it is important to specify whether the morphological data should be corrected for acquisition bias or not (Mk or Mkv options). However, in RAxML and MrBayes, this bias is corrected automatically by addition of ``dummy'' characters.

The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal. Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data, therefore we made the decision to describe the model for morphological characters evolution as the M\textit{k} model (an equal rate Q-matrix with \textit{k} character states where the rate an be equal to 0). RAxML and MrBayes did however corrected for acquisition bias as implemented in their algorithm for treating morphological data.

However, following the reviewer comment, we changed the designation of the model from M\textit{k} to M\textit{kv} to explicitly underline that we corrected for acquisition bias.

%The Mkv model assumes that all the characters are variable, whereas the Mk model is less constrained - some characters can be invariable and others variable. Thus the Mkv model is actually just a special case of the Mk model.

%The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal. Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data %

%However, this is not an issue because we implement opur analyses in RAxML and MrBayes. These programs implement the Mk do not distinguish between both models since the Mkv model is just a particular case of the Mk model and are not distinguished either in the original publication that we cite in the manuscript (Lewis 2001). We've changed the mentions of the M\textit{k} model to M\textit{kv} model throughout our revision.

@@

\textcolor{blue}{D-Many of the equations are incorrect.}

See our detailed response below.

%-------------

\textcolor{blue}{E-If there was a mention of the data being deposited, I missed it. I apologize if I did miss it.}

@@

%-------------

\textcolor{blue}{-------------------------}

\textcolor{blue}{Some specific concerns:}

\textcolor{blue}{1-line 25 and elsewhere: incorrect smart quotes in ``best", ``classical"...}

We fixed the smart quotes in our revision.

@@

%-------------

\textcolor{blue}{2-line 88 ``used it to infer a matrix" $->$ ``used it to simulate a matrix". ``inference" is used in place of simulation elsewhere.}

@@

We changed ``infer" into ``simulate" both at line 88 and in the Figure 1 caption.

%-------------

\textcolor{blue}{3-line 112: does diversitree just stop simulating when it reaches the correct number of tips, or does it deal with the biases that can result from doing that? (Stadler 2011 is the ref for a discussion of this issues, I believe)}

@@

See our response to the comment ``The procedure for generating a tree with fossils is a valid procedure." below.

%-------------

\textcolor{blue}{4-line 113: ``rates from a uniform distribution" with what upper and lower bounds?}

@@

Both birth ($\lambda$) and death ($\mu$) parameters are probabilities bounded between 0 and 1. We have added this to the new version of the manuscript:

%-------------

\textcolor{blue}{5-line 118: with 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}

@@

Lines @@@112 to @@@114 in our revision:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
[We] select only trees with 25 living and 25 fossil taxa. The fossil taxa were considered as unique tips at the end of extinct lineages. % correct in ms
\end{minipage}

%-------------

\textcolor{blue}{6-line 110 - line 120. The procedure for generating a tree with fossils is a valid procedure. But it is idiosyncratic, which will make these results harder to relate to other work. Easy of comparability with other work is one of the advantages of using formally characterized models. In this case, I think that using the TreeSim package http://cran.r-project.org/web/packages/TreeSim/TreeSim.pdf would have led to a study with higher impact. This part of the methods section would be reduced to stating the parameter values chosen, and other workers could study similar tree generation simulations. As it stands, it is unlikely that other workers will every generate trees in precisely the way that the authors do. Thus, there will always be a ``however, Guillerme and Cooper generated their trees under a different model..." caveat necessary when discussing this work in the context of other work.}

We do not agree that the procedure is idiosyncratic. \texttt{diversitree} and \texttt{TreeSim} both generate birth-death trees in the same way (although \texttt{diversitree} is slightly quicker) and are well curated and maintained. We chose to use the \texttt{diversitree} package because we are more familiar with it. We could also have used a number of other packages including \texttt{TESS}. We are also by no means the only people using the \texttt{diversitree} package - the paper associated with it (FitzJohn 2012, Methods in Ecology and Evolution) has been cited three times more (155 cites on Google Scholar) than the publication associated with the \texttt{TreeSim} package (Stadler 2011, Systematic Biology - 55 cites on Google Scholar).

We also disagree that using \texttt{TreeSim} would reduce the amount of explanation required. We believe that carefully outlining our methods will help people repeat our analyses. Thus we would describe any \texttt{TreeSim} analysis in just as much detail.
Finally of course no one can repeat our analyses precisely as we used random parameters to avoid biasing the analyses towards a particular type of tree. However, we do provide all these trees in the supporting information (``Data Reproducibility and Availability" section; \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} link (goo.gl/h5Kxb5)) so future workers can fully repeat our analyses. Thus we feel that repeating the entire set of simulations (1.5 CPU centuries worth) using \texttt{TreeSim} is not necessary and would not influence our results or their utility to future workers.

@@

%-------------

\textcolor{blue}{7-line 126 ``random base frequencies" is too vague. Dirichlet(1, 1, 1, 1) perhaps?}

@@

Lines @@@112 to @@@114 in our revision:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
The matrix [...] was generated [...] using the HKY model with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1)[...]
\end{minipage}

%-------------

\textcolor{blue}{8-Do the simulations apply ascertainment biases as described by Lewis' Mkv (2001) model? It seems like they should... At minimum, the point should be discussed.}

See our response to comment C above.
% NEED to fix!

@@

%We are not entirely sure if the reviewers refers to a sampling bias or to the problem link to constant characters described in Lewis (2001) (acquisition bias). If the reviewers are referring to the second (acquisition bias), both phylogenetic inference software we used (RAxML and MrBayes) are correcting for the this bias (-K MK  see Supplementary information 1 for the full list of commands). Also, the effect of the acquisition bias is to inflate branch length which is an aspect of the phylogeny we didn't explored in this study (focusing on topology). See response below to the comment ``\textit{The use of Mk rather than Mkv makes the study less relevant to real data analyses.}".

%-------------

\textcolor{blue}{9-line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking. It makes a difference. If you mask out $M_F$ percentage of cells, and then mask $M_C$ percentage of columns, it is possible (indeed very probable) that the percentage of missing data in fossil characters that remain will not be $M_F$ after some of the morphological columns are removed. If you mask by $M_C$ first, and then $M_F$, the resulting matrices will always have $M_F$ percentage missing data (modulo rounding error) in the simulated realizations.}

@@

\textcolor{blue}{10-Header for 3. should be ``building phylogenetic estimates" or ``estimating phylogenies" because the previous section described how the phylogenies and character data were ``built."}

We changed the Header 3 to ``Estimating phylogenies".

%-------------

\textcolor{blue}{11-RAxML searching: it would be nice to see some searches started from the true tree. This is clearly not an option for real data sets, but it is helpful in simulation studies because it lets the reader determine if the results could be an artifact of insufficient tree searching.}

@@

Our simulations were designed to test the effects of missing data in as realistic a manner as possible given our constraints. Thus, we do not feel that this additional analysis would add much to the paper, especially since it is already rather long and complex.

%-------------

\textcolor{blue}{12-Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}

We implemented this step of keeping trees with a minimum total median bootstrap to make the topological comparisons more robust. If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the fact that low supported nodes can be just random. We've added the following justification in our revision:

@@

Lines @@@112 to @@@114 in our revision:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
We repeated this selection until we obtained 50 sets of simulations (i.e. 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50). This step was implemented to make sure that the differences we observed in topologies (see below) where due to the amount of missing data for each parameters ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
\end{minipage}

%-------------

\textcolor{blue}{13-MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be available in real data analysis in which one does not know the true values) for the Bayesian method. This makes the studies report of superior behavior of Bayesian methods less persuasive.}

See our response to comment B above.

@@

%-------------

\textcolor{blue}{14-line 254: RF measures the difference between the number of clades and the twice number of shared clades across two trees. The statement that it ``measures the number of shared clades across two trees" should be revised.}

@@

Lines @@@112 to @@@114 in our revision:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
The Robinson-Foulds distance, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees.
\end{minipage}

%-------------

\textcolor{blue}{15-line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}

We thanks the reviewer for pointing at this error and we changed the following sentence:

Lines @@@112 to @@@114 in our revision:

@@

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
\end{minipage}

%%%%%%%%%%%%%%%%%
%CHECK in Soltis!
%%%%%%%%%%%%%%%%%

%-------------

\textcolor{blue}{16-The ``scaling" to produce the Normalized RF ``distance" is confusing. It is also inverting the sign of the distance, so it has become a measure of similarity.}

We agree that this is confusing and thank the reviewer for pointing this inconsistency out. We changed the mentions of ``Normalised Robinson-Foulds distance'' and ``Normalised Triplets distance'' to ``Normalised Robinson-Foulds metric'' and ``Normalised Triplets metric'' throughout the our revision. Additionally, we clarified the scaling method as follows:

Lines @@@112 to @@@114 in our revision:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
We normalised this metric following Bogdanowicz \textit{et al.} (2012)'s Normalised Tree Similarity (NTS) method. This methods scales %not sure if scale here?
any tree comparison metric using the mean distance between 1000 random trees (see Supporting Information S2: Tree Comparisons for the calculation details). This method is a generalisation of the topological accuracy method (Price \textit{et al.} 2010) allowing to compare topological differences between any tree with any tree comparison metric. In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
\end{minipage}
% Note that NRF distance is actually a similarity ...

@@

%-------------

\textcolor{blue}{17-line 275: why isn't the upper bound of the triples distance n choose 3 (instead of n choose 4). I think you got the formula from an unrooted quartet distance.}

@@

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
It is bounded between zero when the two trees are identical and $\binom{n}{3}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
\end{minipage}

%%%%%%%%%%%%%%%%%
%CHEK in Soltis!
%%%%%%%%%%%%%%%%%

%-------------

\textcolor{blue}{18-Using the BC stat may be similar in spirit to using a t-test, but it is not equivalent.}

@@

Lines @@@112 to @@@114 in our revision:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
Note that this is similar to performing a two-sided t-test for comparing two distribution.
\end{minipage}

%-------------

\textcolor{blue}{19-Figure 3: B is confusing. If both curves are probability distributions, then they both have to integrate to 1. So B(y) can't cover just a subset of the area covered by B(x).}

We thank the reviewer for pointing out this error. We fixed the shape of the distribution in Figure 3-B.

@@

%-------------

\textcolor{blue}{20-``Combined effect of missing data parameters" section ANOVA or multiple regression are the appropriate analyses to tease apart multiple interacting factors. Here the authors opt to describe results without statistical tests, which is disappointing.}

@@

\textcolor{blue}{Supplement 2:}
\textcolor{blue}{22-more confusion about the min RF here. Also note that if you are counting the entire leaf set as a clade (e.g the star tree having N=1), then the number of clades in rooted binary tree is n-1 (not n-2). Equation 3 has an extra ``-2" As stated it implies that the max RF distance for n=3 is 0, when the correct answer is 2.}

%check

%-------------

\textcolor{blue}{23-The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}

@@

%not in this paper

%-------------

\textcolor{blue}{24-The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees. I'm not sure that it makes much difference, but it is odd.}

@@

We cited the reference (Bogdanowicz 2012) of the method.
%explain in a bit more detail

%-------------

\textcolor{blue}{25-The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0. This is not true of either distance metric here. So the real range of values is (mean-max)/mean for both the RF and triples distance.}

As clarified in our revision (lines @@@), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.

@@

%-------------

\textcolor{blue}{26-Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}

@@

\textcolor{blue}{27-Equation (6) is not correct. You shouldn't set the summation sign as the left hand side of an equation because the notation has a standard definition. More importantly, the $\#$ of triples is the number of ways of drawing 3 (unordered) leaves from the leaf set. So it is n choose 3 not n choose 4.}

%check

%-------------

### 2015-04-28 Manuscript/PLoSONEresponse/TEM_PLoS_response.tex
Fixed some comments (incl. figures)

\paragraph*{Robinson-Foulds distance}
The Robinson-Foulds distance \cite{RF1981}, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees. The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see ~\nameref{S2_Appendix2} for calculation details). This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees. This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity. We normalised this metric following Bogdanowicz \textit{et al.} \cite{Bogdanowicz2012}'s Normalised Tree Similarity (NTS) method. This methods scales %not sure if scale here?
any tree comparison metric using the mean distance between 1000 random trees (see ~\nameref{S2_Appendix2} for the calculation details). This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric. In practice when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.

\paragraph*{Triplets distance}
The Triplets distance \cite{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees \cite{critchlowthe1996} (see ~\nameref{S2_Appendix2} for calculation details). This metric measures the position of each taxon and clade in relation to its closest neighbours. It is bounded between zero when the two trees are identical and $\binom{n}{3}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees. Therefore this metric sensitive to the conservation of wildcard taxa. We normalised this metric in the same way as for the Robinson-Foulds distance resulting in the Normalised Triplets distance.

\paragraph*{Paired tree comparisons}
\label{tree_comparisons}
For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance for each tree inference method. Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the ``best" trees, and then did the same for the 125 ``missing data" trees (resulting in 1000 ``best" trees and 125$\times$1000 ``missing data" trees). 
For a given set of 1000 ``missing data" trees and the 1000 ``best" trees, we sampled one ``missing data" tree and one ``best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above. We repeated this 1000 times for each set of ``missing data" trees resulting in 125$\times$1000 values for each metric. We repeated all the paired tree comparisons described above for each of the 50 simulation runs. We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v. 3.1 \cite{hdrcde}.

@@

We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \cite{Bhattacharyya}. The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see ~\nameref{S2_Appendix2} for calculation details). Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means. To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc. (see Fig.~\ref{Fig_Bhattacharyya_Coefficients1} for more details). This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells). We performed this procedure separately for each tree inference and uncertainty method. When two combinations of missing data parameters have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one. Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.

\begin{figure}[h]
\caption{{\bf Bhattacharyya Coefficient calculation outline 1.}
 A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B. Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
\label{Fig_Bhattacharyya_Coefficients1}
\end{figure}

To assess the effect of the different tree inference and uncertainty methods (i.e. Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the ``best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods, i.e. Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc. (see Fig.~\ref{Fig_Bhattacharyya_Coefficients2} for more details). Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons. When two tree inference or uncertainty methods have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one. Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.

\begin{figure}[h]
\caption{{\bf Bhattacharyya Coefficient calculation outline 2.}
 A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g. Maximum Likelihood or Bayesian). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$. Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
\label{Fig_Bhattacharyya_Coefficients2} 
\end{figure}

@@

\section*{Results}
As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance). However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).

\begin{figure}[]
\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).}
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
\label{Fig_Results-permeth_perparam} 
\end{figure}

>>>>

\paragraph*{Robinson-Foulds distance}
The Robinson-Foulds distance \cite{RF1981}, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees. The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see ~\nameref{S2_Appendix2} for calculation details). This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees. This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity. We normalised this distance following Bogdanowicz \textit{et al.} \cite{Bogdanowicz2012}'s Normalised Tree Similarity (NTS) method. This methods scales %not sure if scale here?
any tree comparison metric using the mean distance between 1000 random trees (see ~\nameref{S2_Appendix2} for the calculation details). This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric. In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.

\paragraph*{Triplets distance}
The Triplets distance \cite{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees \cite{critchlowthe1996} (see ~\nameref{S2_Appendix2} for calculation details). This metric measures the position of each taxon and clade in relation to its closest neighbours. It is bounded between zero when the two trees are identical and $\binom{n}{3}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees. Therefore this metric sensitive to the conservation of wildcard taxa. We normalised this metric in the same way as for the Robinson-Foulds distance resulting in the Normalised Triplets metric.

\paragraph*{Paired tree comparisons}
\label{tree_comparisons}
For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robinson-Foulds metrics and 125 Normalised Triplets metric for each tree inference method. Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the ``best" trees, and then did the same for the 125 ``missing data" trees (resulting in 1000 ``best" trees and 125$\times$1000 ``missing data" trees). 
For a given set of 1000 ``missing data" trees and the 1000 ``best" trees, we sampled one ``missing data" tree and one ``best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above. We repeated this 1000 times for each set of ``missing data" trees resulting in 125$\times$1000 values for each metric. We repeated all the paired tree comparisons described above for each of the 50 simulation runs. We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v. 3.1 \cite{hdrcde}.

@@

We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \cite{Bhattacharyya}. The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see ~\nameref{S2_Appendix2} for calculation details). Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means. To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc. (see Fig.~\ref{Fig_Bhattacharyya_Coefficients1} for more details). This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells). We performed this procedure separately for each tree inference and uncertainty method. When two combinations of missing data parameters have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one. Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.

\begin{figure}[h]
\caption{{\bf Bhattacharyya Coefficient calculation outline 1.}
 A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B. Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
\label{Fig_Bhattacharyya_Coefficients1}
\end{figure}

To assess the effect of the different tree inference and uncertainty methods (i.e. Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the ``best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of tree inference and uncertainty methods, i.e. Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc. (see Fig.~\ref{Fig_Bhattacharyya_Coefficients2} for more details). Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons. When two tree inference or uncertainty methods have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one. Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.

\begin{figure}[h]
\caption{{\bf Bhattacharyya Coefficient calculation outline 2.}
 A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g. Maximum Likelihood or Bayesian). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$. Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
\label{Fig_Bhattacharyya_Coefficients2} 
\end{figure}

@@

\section*{Results}
As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric). However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).

\begin{figure}[]
\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).}
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
\label{Fig_Results-permeth_perparam} 
\end{figure}

\begin{figure}[]
\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).}
 The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.} 
\label{Fig_Results-global_perparam}
\end{figure}
